---
title: "Sales Analysis Report"
author: "Jivesh Gawde"
date: "`2025-03-02`"
output: html_document
---

```{r setup, include=FALSE}

source("analysis_retail.R")

# Set global plot options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# **Introduction**  
This report explores customer spending habits and sales trends across different product categories.  
We aim to:  
- Identify top-selling categories.  
- Find high-value customers.  
- Analyze price distributions.  
- Examine sales trends over time.  
- Understand regional and mall-wise performance.  

---

# **Data Summary & Overview**  

```{r summary, echo=TRUE}
summary(sales)
```

---

# **Sales Quantity Distribution**  

- **Minimum quantity sold:** 1  
- **Maximum quantity sold:** 5  
- **Most common (median) quantity:** 3  
- **Average quantity per transaction:** ~3  
- Customers typically buy in small quantities (1-5 items).  
- **Highest sales volume:** **CLOTHING** (62,366 units)  
- **Lowest sales volume:** **TECHNOLOGY** (8,989 units)  


```{r category_wise_quantity}
category_wise_quantity
```

---

# **Revenue Distribution Across Categories**  

```{r category_wise_revenue}
category_wise_revenue
```

---

# **Pricing Patterns**  

- **Minimum price:** $5.23  
- **Maximum price:** $5,250.00  
- **Median price:** $203.30  
- **Mean price:** $687.57  
- The **average price is significantly higher than the median**, indicating some **high-priced outliers** are skewing the data.  
- We will identify these high-priced outliers using a boxplot.  


```{r product_distribution}
product_distribution
```

---

# **Sales Trend Over Time**  

- **Earliest recorded sale:** January 1, 2021  
- **Latest recorded sale:** March 8, 2023  
- The dataset spans **26 months (just over 2 years).**  
- We analyze monthly sales trends to check for **seasonal patterns** and overall growth.  

```{r monthly_sales}
monthly_sales
```

---

# **High-Value Customers**  

- We identify top-spending customers in each category.  
- **Customers who buy Technology tend to be high-value customers** despite fewer purchases.  
- **Clothing has the highest revenue due to higher sales volume.**  

```{r value_customer}
value_customer
```

---

# **Shopping Mall & Regional Insights**  

- Sales are distributed across **multiple malls and regions.**  
- Some malls/regions might be **outperforming others.**  
- We check **total sales by mall & region** to identify top and low-performing locations.  

```{r shop_sales}
shop_sales
```

---

## **Conclusion**  

- **Clothing is the highest-selling category.**  
- **Technology has lower sales volume but high-value customers.**  
- **Price distribution is skewed due to a few expensive products.**  
- **Sales trends show seasonality and a recent drop (likely due to incomplete data).**  
- **Mall & regional performance needs deeper analysis to identify the best locations.**  

---

### **End of Report**  

---

